\documentclass{article}


% ready for submission
\usepackage[final]{hlt_project_template}
\usepackage{natbib}
\usepackage{float}
\usepackage{graphicx}
\bibliographystyle{plainnat}



\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{siunitx}    % for \SI and \num macros (percentages, large numbers)
\usepackage{tikz}
\usetikzlibrary{positioning}


\title{Trip Advisor Reviews Sentiment Analysis using RNN and Transformers} 


\author{%
  Salvatore Mastrangelo\\
  \texttt{s.mastrangelo4@studenti.unipi.it}
}


\begin{document}

\maketitle


\begin{abstract}
This study investigates document-level sentiment classification of hotel reviews 
by contrasting a lightweight Long Short-Term Memory Recurrent Neural Network (LSTM-RNN) 
with a pretrained Transformer (RoBERTa-base) fine-tuned through Low-Rank Adaptation (LoRA). 
The analysis is conducted on 18'273 English hotel reviews trimmed to 256 tokens and 
split 72.3 / 12.7 / 15 \% for training, validation and testing, respectively, in a strongly 
imbalanced five-star rating distribution.

The LSTM comprises an embedding layer, a single LSTM layer with mean-pooled hidden states, 
and a soft-max classifier, achieving convergence below 240 s on a consumer GPU. 
A RoBERTa based model was fine-tuned with FP16 precision and a LoRA adapter, thet reduced 
trainable parameters to 1.9 \% of the backbone, enabling single-GPU training.

On the native dataset, RoBERTa outperformed the LSTM with 67 \% accuracy / weighted 
F1 = 0.67 in the five-class task and 88 \% / 0.88 in the ternary task, compared with 58 \% / 0.58 and 84 \% / 0.83 for the LSTM, respectively. 
Transformer models shows great capability in classifying minority ratings (1–3 stars). 

These findings underscore three insights: (i) pre-trained attention models confer 
a clear advantage on imbalanced, ambiguous opinion data; (ii) conventional RNNs remain 
competitive under balanced conditions and deliver rapid training; (iii) LoRA enables 
cost-effective fine-tuning of large encoders. The work thus provides a reproducible 
benchmark and practical guidelines for choosing architectures under resource and 
data-distribution constraints.

% This report presents a comprehensive analysis of sentiment classification on Trip Advisor reviews using 
% Recurrent Neural Networks (RNNs) and Transformers. The study explores the effectiveness of these models 
% in understanding and predicting sentiment from textual data. A detailed methodology, experimental setup, 
% and results are provided, highlighting the strengths and weaknesses of each approach. In particular,
% given a rather small dataset of Trip Advisor reviews, the aim is to classify the sentiment of each review into both
% stars classes (1 to 5 stars) and ternary classes (positive, negative, neutral). RNN models are implemented
% using Long Short-Term Memory (LSTM), while Transformers are implemented using the BERT architecture, in
% particular the RoBERTa model. The results show that while Transformers require much less epochs to 
% converge, RNNs are able to achieve better performance in terms of accuracy and F1 score on such
% small dataset. On the other hand, Transformers show much better capabilities in terms of ambiguous reviews, 
% achieving a better F1 score and accuracy in the middle part of the scale (3-4 stars). 
\end{abstract}

\input{chapters/01_introduction}
\input{chapters/02_background}
\input{chapters/03_method}
\input{chapters/04_experimental_analysis}
\input{chapters/05_discussion}
\input{chapters/06_conclusions}

\bibliography{bib.bib}

\small

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}